\section{Installation}
\label{sec:installation}

In this section, we will describe the various ways that one can install the 
\thispackage package.


\subsection{Important Notes}

It is possible for \PAPI and/or \thispackage to build, but still have no access
to any counter information.  For example, the package will build on the popular 
Arm architecture device 
Raspberry Pi\footnote{\url{http://www.raspberrypi.org/}};
however, it does not have any hardware counters, and as such, \thispackage 
effectively can not be used.  If the package installs without error, you can 
see how many hardware counters you have available by calling:
\begin{lstlisting}[language=rr]
library(pbdPAPI)
system.ncounters()
\end{lstlisting}

However, this too has some caveats.  Those with Intel Sandy Bridge and/or Intel 
Ivy Bridge architectures should be aware that flops counts are unreliable on 
these platforms.  This is a problem with the hardware returning incorrect 
values, not with \PAPI or \thispackage.  For more details, see: 
\href{http://icl.cs.utk.edu/projects/papi/wiki/PAPITopics:SandyFlops}%
{the \PAPI documentation} on this matter.


Additionally, not all platforms support all counters.  For the high-level 
interface, this simply means that some things you may wish to try simply can 
not physically work, and as such will return an error.  

For a detailed list of all counters and whether or not your particular platform 
supports them, simply call 
\begin{lstlisting}[language=rr]
library(pbdPAPI)
papi.avail()
\end{lstlisting}





\subsection{Installing WITHOUT a System Installation of PAPI}

This is the default method of installation.  Here, the \PAPI library will  
automatically be built first as a static library, and then the \thispackage 
package will be built and linked against that static library.  All of this is 
handled completely transparently, and should only go wrong if your system is not 
supported by \PAPI.  This is the simplest approach, and should cover most users. 
 Simply build the package as you would any other:
\begin{Command}
R CMD INSTALL pbdPAPI_0.1-0.tar.gz
\end{Command}
and using the \pkg{devtools} package:
\begin{lstlisting}[language=rr]
library(devtools)
install_github(username="wrathematics", repo="pbdPAPI")
\end{lstlisting}




\subsection{Installing WITH an Existing System Installation of PAPI}

If you already have a system installation of \PAPI available, it makes more  
sense to link with that existing library.  The one catch is that the static 
library \emph{must} have been compiled with \code{-fPIC}, which is non-standard. 
 To build an external \PAPI library in this way, you should do so by first 
setting:
\begin{Command}
export CC="${CC} -fPIC"
\end{Command}
Assuming that \code{CC} is set; if not, you can use \code{cc} in the right hand 
side.

To link with an external installation of \PAPI, from the command line, execute:
\begin{Command}
R CMD INSTALL pbdPAPI_0.1-0.tar.gz \ 
    --configure-args="--enable-system-papi \ 
    --with-papi-home=location/of/PAPI/install"
\end{Command}
and using the \pkg{devtools} package:
\begin{lstlisting}
library(devtools)
install_github(username="wrathematics", repo="pbdPAPI", 
    args="--configure-args='--enable-system-papi 
          --with-papi-home=location/of/PAPI/install'")
\end{lstlisting}
